---
title: "Reproducible Research: Peer Assessment 1"
author: "rodrigo zuniga"
date: "09/03/2020"
output: html_document
---
## Loading and preprocessing the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
rm(list = ls())
setwd("C:\\Users\\rodri\\Dropbox\\Education\\R\\RClass5\\repdata_data_activity")
activity<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
totday<-tapply(activity$steps,FUN=sum,na.rm=T,INDEX=activity$date)
hist(totday,main="Daily Steps Histogram",xlab = "Daily Steps")
meandailysteps<-mean(totday)
mediandailysteps<-median(totday)
```
- Mean number of daily steps      : `r round(meandailysteps)`  
- Median number of daily steps    : `r mediandailysteps`  

## What is the average daily activity pattern?
```{r}
meanstepsbyinterval<-tapply(activity$steps,FUN=mean,na.rm=T,INDEX=activity$interval)
plot(row.names(meanstepsbyinterval),meanstepsbyinterval,
     ylab = "Avg # Steps",xlab="Time Interval",main="Avg #Steps by Time Interval",
     type="l")
maxsteps=max(meanstepsbyinterval)
maxinterval<-meanstepsbyinterval[meanstepsbyinterval==maxsteps]
maxintervaltime<-rownames(meanstepsbyinterval)[which(meanstepsbyinterval==maxsteps)]
```

- Maximum number of steps in interval:
        ```r round(maxinterval)```
- Time with maximum number of steps:
        ```r maxintervaltime```

## Imputing missing values
```{r}
medianstepsbyinterval<-tapply(activity$steps,FUN=median,na.rm=T,INDEX=activity$interval)

medianstepsbyinterval2<-data.frame(mediansteps=medianstepsbyinterval,interval=as.numeric(row.names(medianstepsbyinterval)))

activity2<-dplyr::inner_join(activity,medianstepsbyinterval2,by="interval")

activity2$steps2<-(ifelse(is.na(activity2$steps),activity2$mediansteps,activity2$steps))

totday2<-tapply(activity2$steps2,FUN=sum,na.rm=T,INDEX=activity2$date)

hist(totday2,main="Daily Steps Histogram \n w/ imputed values",xlab = "Daily Steps")

oldmean<-mean(totday)
newmean<-mean(totday2)

oldmedian<-median(totday)
newmedian<-median(totday2)
```

- mean before imputation: ```r round(oldmean)```
- mean after imputation: ```r round(newmean)```

- median before imputation: ```r oldmedian```
- median after imputation: ```r newmedian```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
activity2$weekpart<-factor(ifelse(weekdays(as.Date(activity2$date))%in%c("Saturday","Sunday"),"weekend"
                                  ,"weekday"))
weekpartsteps<-aggregate(activity2$steps2,FUN=mean,by = list(interval=activity2$interval,weekpart=activity2$weekpart))
weekpartsteps<-dplyr::rename(weekpartsteps,steps=x)

xyplot(steps~interval|weekpart,data=weekpartsteps,type="l",layout=c(1,2),xlab = "Interval",ylab="Number of steps")
```
